---
title: "Factorial Experiments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Factorial Experiments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This vignette shows how to build simulation definitions programmatically.

## Parameters

A top-level parameter changes a value outside a named LPJ-GUESS block.

```{r}
top_level_parameter("npatch", 10)
top_level_parameter("wateruptake", "rootdist")
```

A block parameter changes a value inside a named block, such as a PFT block.

```{r}
block_parameter("pft", "MRS", "sla", 26)
block_parameter("pft", "MRS", "include", TRUE)
```

## Simulations

Pass parameter changes to `simulation()`.

```{r}
simulation(
    "rootdist_sla_26",
    top_level_parameter("wateruptake", "rootdist"),
    block_parameter("pft", "MRS", "sla", 26)
)
```

You can also pass a list of changes.

```{r}
changes <- list(
    top_level_parameter("wateruptake", "rootdist"),
    block_parameter("pft", "MRS", "sla", 26)
)

simulation("rootdist_sla_26", changes)
```

## Parameter Grids

R's `expand.grid()` is convenient when you want one simulation for each
combination of parameter values.

```{r}
grid <- expand.grid(
    wateruptake = c("wcont", "rootdist"),
    sla = c(26, 39),
    stringsAsFactors = FALSE
)

simulations <- lapply(seq_len(nrow(grid)), function(i) {
    values <- grid[i, ]
    simulation(
        sprintf("wu_%s_sla_%s", values$wateruptake, values$sla),
        top_level_parameter("wateruptake", values$wateruptake),
        block_parameter("pft", "MRS", "sla", values$sla)
    )
})
```

The resulting `simulations` list can be passed directly to
`run_simulations()`.

```{r}
result <- run_simulations(
    settings,
    simulations,
    instruction_files = "/path/to/site.ins",
    pfts = "MRS"
)
```

Simulation names are used in output discovery and should be unique within an
experiment whenever possible.
