---
title: "Reading Outputs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading Outputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Completed runs are catalogued below the output directory configured in
`run_settings()` or `run_settings_local()`.

## Read From a Result

The result returned by `run_simulations()` knows where the run wrote outputs.

```{r}
result <- run_simulations(settings, simulations, "/path/to/site.ins")
lai <- read_output(result, "lai")
```

`read_output()` returns the LPJ-GUESS output columns in their original wide
form, with a leading `simulation` column.

## Read From a Previous Run

You can also discover and read outputs from an existing output directory.

```{r}
runs <- list_simulations("/scratch/me/lpjguess-runs")
outputs <- list_outputs(runs)
logs <- read_logs(runs)

lai <- read_output(runs, "lai")
```

`list_simulations()` returns one row per completed simulation job. Simulation
names are the primary user-facing identifier.

```{r}
baseline <- read_output(runs, "lai", simulations = "baseline")
```

If the same simulation name was run for multiple base instruction files, filter
the simulation table first or provide `base_ins`.

```{r}
site1 <- subset(runs, simulation == "baseline" & basename(base_ins) == "site1.ins")
lai_site1 <- read_output(site1, "lai")
```

## Output Names

`read_output()` accepts the LPJ-GUESS instruction-file parameter, the concrete
output filename, or the filename stem.

```{r}
read_output(runs, "file_lai")
read_output(runs, "lai.out")
read_output(runs, "lai")
```

For conventional output names, the stem also matches the corresponding
`file_*` parameter.

```{r}
read_output(runs, "dave_lai")
```

If multiple outputs are requested, `read_output()` adds an `output_type` column
so the rows can be distinguished.

```{r}
carbon <- read_output(runs, c("lai", "agpp"))
```

If one requested output does not exist, a warning is emitted and matching
outputs are still returned.

## Table Type

If `data.table` is installed, `read_output()` uses `data.table::fread()` and
returns a `data.table`. This is faster for large LPJ-GUESS output files and
prints compactly at the console.

Without `data.table`, the function falls back to base R and returns a
`data.frame`.

## Metadata Columns

The default result includes only the identifiers most users need. Use
`id_cols = "all"` to keep paths and internal catalog identifiers.

```{r}
lai <- read_output(runs, "lai", id_cols = "all")
```

Use `id_cols = FALSE` to read only the raw LPJ-GUESS columns.

```{r}
raw_lai <- read_output(runs, "lai", id_cols = FALSE)
```

## Logs

LPJ-GUESS writes `guess.log` next to each generated instruction file and
manifest. Use `list_logs()` to find the log paths, or `read_logs()` to read
them into a table.

```{r}
logs <- list_logs(runs)
log_lines <- read_logs(runs)
```

The default table has `simulation`, `line`, and `text` columns. Use
`id_cols = "all"` to include paths and catalog metadata.

```{r}
log_lines <- read_logs(runs, simulations = "baseline")
raw_logs <- read_logs(runs, combine = FALSE)
```
