---
title: "Running Simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running Simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This vignette covers the main ways to run LPJ-GUESS simulations from R.

## Configuration

Use `run_settings_local()` when LPJ-GUESS should run directly on the current
machine.

```{r}
library(lpjguessRunner)

settings <- run_settings_local(
    guess_path = "/path/to/guess/executable",
    output_directory = "/desired/path/to/output",
    input_module = "nc",
    cpu_count = 4,
    job_name = "experiment"
)
```

The output directory is where generated instruction files, manifests, and
LPJ-GUESS outputs are written.

The guess executable will typically be found in the build directory of an
LPJ-GUESS checkout, e.g. `~/lpj-guess/build/guess`. On Windows, it will be
`guesscmd.exe` (e.g. `C:\lpj-guess\build\guesscmd.exe`).

## Running Simulations

A simulation is a named set of changes applied to each instruction file.

```{r}
simulations <- list(
    simulation("baseline"),
    simulation(
        "high_sla",
        block_parameter("pft", "MRS", "sla", 39)
    )
)
```

Run the experiment synchronously with `run_simulations()`.

```{r}
result <- run_simulations(
    settings = settings,
    simulations = simulations,
    instruction_files = c("/path/to/site1.ins", "/path/to/site2.ins"),
    pfts = "MRS"
)
```

LPJ-GUESS standard output/error messages are captured but not printed during
successful runs. If a run fails, the captured output is printed before
`run_simulations()` raises an R error.

## Output Streaming

Set `progress = NULL` to suppress the default progress line, or pass a function
to handle progress yourself.

```{r}
result <- run_simulations(
    settings,
    simulations,
    "/path/to/site.ins",
    progress = function(event) {
        message(sprintf("%0.1f%% complete", event$percent))
    }
)
```

Pass an `output` callback if you want to stream model output during the run.

```{r}
result <- run_simulations(
    settings,
    simulations,
    "/path/to/site.ins",
    output = function(event) {
        cat(sprintf("[%s][%s] %s\n", event$job, event$stream, event$text))
    }
)
```

## Existing Outputs

`existing_output_policy` controls how managed output directories are handled
when a run is repeated.

```{r}
result <- run_simulations(
    settings,
    simulations,
    "/path/to/site.ins",
    existing_output_policy = c("clean_managed", "prune_stale")
)
```

Common policies are:

- `"clean_managed"`: remove old files for simulations being regenerated.
- `"prune_stale"`: remove managed simulations from previous runs that are not
  part of the current experiment.
- `"preserve"`: leave existing outputs untouched.
- `"fail"`: abort if existing output directories are found.

## PBS Runs

Use `run_settings()` with `run_local = FALSE` for PBS-backed execution.

```{r}
settings <- run_settings(
    guess_path = "/apps/lpj-guess/guess",
    output_directory = "/output/path",
    input_module = "nc",
    cpu_count = 16,
    job_name = "experiment",
    run_local = FALSE,
    walltime = "12:00:00",
    memory = 32,
    queue = "normal",
    project = "project-code"
)
```

## Asynchronous Runs

Use `run_simulations_async()` if R needs to remain responsive while the runner
is active.

```{r}
handle <- run_simulations_async(settings, simulations, "/path/to/site.ins")

while (poll_run(handle, timeout = 100)) {
    # Do other R work between polls.
}

result <- wait_run(handle)
```

Cancel a running experiment with `cancel_run(handle)`.
